### 2019-09-06 results/yogo_empirical_results_2019-08-21 17:08:23.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Aug 21 17:08:23 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrrlrrrrrl}
  \hline
name\_list & var\_list & obs\_list & delta\_list & rho\_ci\_list & df\_gls\_list & lags\_list & t\_stat\_list & as.numeric.pretest\_list. & beta\_list & beta\_ci\_scaled\_list \\ 
  \hline
CRSP\_M & lep & 913 & -0.99 & [0.985,1] & -1.859 & 1 & 2.66 & 0 & 0.013 & [0.001,0.018] \\ 
  CRSP\_M & ldp & 913 & -0.95 & [0.99,1.001] & -1.433 & 1 & 1.70 & 0 & 0.008 & [-0.005,0.01] \\ 
   \hline
\end{tabular}
\end{table}