### 2019-09-06 results/yogo_empirical_results_2019-08-21 17:27:00.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Aug 21 17:27:00 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrrlrrrrrl}
  \hline
name\_list & var\_list & obs\_list & delta\_list & rho\_ci\_list & df\_gls\_list & lags\_list & t\_stat\_list & as.numeric.pretest\_list. & beta\_list & beta\_ci\_scaled\_list \\ 
  \hline
SP\_A & lep & 78 & -0.97 & [0.796,0.975] & -2.091 & 1 & 2.70 & 0 & 0.155 & [0.02,0.244] \\ 
  SP\_A & ldp & 78 & -0.87 & [0.84,0.978] & -1.820 & 1 & 2.05 & 0 & 0.106 & [-0.009,0.205] \\ 
   \hline
\end{tabular}
\end{table}