### 2019-09-06 results/yogo_empirical_results_2019-08-21 18:31:15.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Wed Aug 21 18:31:15 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrrlrrrrrl}
  \hline
name\_list & var\_list & obs\_list & delta\_list & rho\_ci\_list & df\_gls\_list & lags\_list & t\_stat\_list & as.numeric.pretest\_list. & beta\_list & beta\_ci\_scaled\_list \\ 
  \hline
SP\_A & lep & 115 & -0.96 & [0.69,0.873] & -3.519 & 1 & 3.32 & 0 & 0.189 & [0.092,0.323] \\ 
  SP\_A & ldp & 115 & -0.84 & [0.83,0.949] & -2.444 & 3 & 2.22 & 0 & 0.130 & [-0.024,0.227] \\ 
   \hline
\end{tabular}
\end{table}