### 2019-09-07 results/yogo_empirical_results_A_2019-09-07 19:30:31.tex
ADD empirical results rep

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Sat Sep  7 19:30:31 2019
\begin{table}[ht]
\centering
\begin{tabular}{llrrlrrrrrl}
  \hline
name\_list & var\_list & obs\_list & delta\_list & rho\_ci\_list & df\_gls\_list & lags\_list & t\_stat\_list & as.numeric.pretest\_list. & beta\_list & beta\_ci\_scaled\_list \\ 
  \hline
CRSP\_A & lep & 77 & -0.96 & [0.778,0.96] & -2.229 & 1 & 2.77 & 0 & 0.162 & [0.04,0.273] \\ 
  CRSP\_A & ldp & 77 & -0.72 & [0.926,1.01] & -1.033 & 1 & 2.53 & 0 & 0.158 & [0.013,0.186] \\ 
   \hline
\end{tabular}
\end{table}